dataCleaning

### 2024-03-25 scripts/dataCleaning.qmd
prepared dataCleaning.qmd for new vultureUtils functions

# Load information about capture sites
capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
load(here::here("data", "dataPrep", "roosts0.Rda"))
AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))

removeCaptures <- function(data, roosts, captureSites, distance = 500){
  # Identify the period of time during which the capture sites are open (when we need to do this exclusion)
  start.day <- 01
  start.month <- 08
  end.day <- 30
  end.month <- 11

  # Get roosts that fall in the capture period
  sub.roosts <- roosts %>%
    mutate(start_date = lubridate::dmy(paste(start.day, start.month,
                                      lubridate::year(date), sep = "-")),
           end_date = lubridate::dmy(paste(end.day, end.month, 
                                           lubridate::year(date), sep = "-"))) %>%
    filter(date >= start_date & date <= end_date)

  if(nrow(sub.roosts) > 0){
    # Calculate the roost distance to each of the capture cages. If it is less than 500m, keep that line.
  crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
                 nrow = nrow(sub.roosts), ncol = 2) # roost locs as simple lat/long coords
  distanceMat <- matrix(ncol = nrow(captureSites),
                        nrow = nrow(crds))
  colnames(distanceMat) <-  unique(captureSites$name)

  for(i in 1:nrow(crds)){
    distanceMat[i,] <- round(geosphere::distm(crds[i,], 
                                              captureSites[,c(3,2)]), 2)
    }
  }

  closestCaptureSite <- colnames(distanceMat)[apply(distanceMat, 1, which.min)] # ID of closest capture site
  closestCaptureDist <- apply(distanceMat, 1, min) # distance from closest capture site

  sub.roosts <- cbind(sub.roosts, closestCaptureSite, closestCaptureDist)
  sub.roosts$captured <- ifelse(sub.roosts$closestCaptureDist <= distance, T, F)
  sub.captured.dates <- sub.roosts %>% 
    filter(captured) %>%
    select(Nili_id, date, closestCaptureSite, 
           closestCaptureDist, captured)
  # XXX START HERE

}

>>>>

# Load information about capture sites
capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))

# Full code as it currently exists in the data cleaning pipeline:
# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
start.day <- 01
start.month <- 08
end.day <-  30
end.month <- 11
distance <- 500 # distance, in meters, to calculate from the cage

# Subset the roost dataset with the start and end dates for the capture period
sub.roosts <- roosts0 %>%
  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
                              format="%d-%m-%Y"),
         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
                            format="%d-%m-%Y")) %>%
  filter(date >= start_date & date <= end_date)

unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.

# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat),
               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates

DistanceMat <- matrix(ncol = nrow(capture_sites), nrow = nrow(crds))
colnames(DistanceMat) <- unique(capture_sites$name)

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], capture_sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  #print(i)
}

ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site

sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist)
sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")

sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage

sub.captured.dates <- subset(sub.captured,
                             select = c("Nili_id",
                                        "date",
                                        "ClosestCaptureSite",
                                        "ClosestCaptureDist",
                                        "Captured"))
